---
title: 'Psych 254 W15 PS #1'
author: "Mike Frank"
date: "January 11, 2015"
output: html_document
---

This is problem set #1, in which we hope you will practice the packages tidyr and dplyr. There are some great cheat sheets from [RStudio](http://www.rstudio.com/resources/cheatsheets/).

The data set
------------

This data set comes from a replication of [Janiszewski and Uy (2008)](http://dx.doi.org/10.1111/j.1467-9280.2008.02057.x), who investigated whether the precision of the anchor for a price influences the amount of adjustment.

In the data frame, the `Input.condition` variable represents the experimental condition (under the rounded anchor, the rounded anchor, over the rounded anchor). `Input.price1`, `Input.price2`, and `Input.price3` are the anchors for the `Answer.dog_cost`, `Answer.plasma_cost`, and `Answer.sushi_cost` items.

Preliminaries
-------------

I pretty much always clear the workspace and load the same basic helper functions before starting an analysis.

```{r prelims}
rm(list=ls())
source("../helper/useful.R")
```

Note that I'm using a "relative" path (the "../helper") rather than an absolute path (e.g. "/Users/mcfrank/code/projects/etc..."). The relative path means that someone else can run your code by changing to the right directory, while the absolute path will force someone else to make trivial changes every time they want to run it.

Part 1: Data cleaning
---------------------

The first part of this exercise actually just consists of getting the data in a format usable for analysis. This is not trivial. Let's try it:

```{r data1}
d <- read.csv("../data/janiszewski_rep_exercise.csv")
```

Fine, right? Why can't we go forward with the analysis?

HINT: try computing some summary statistics for the different items. Also, are there any participants that did the task more than once?

Fix the data file so that it looks reasonable.



Part 2: Making these data tidy
------------------------------

Now let's start with the cleaned data, so that we are all beginning from the same place.

```{r data2}
d <- read.csv("../data/janiszewski_rep_cleaned.csv")
```

This data frame is in *wide* format - that means that each row is a participant and there are multiple observations per participant. This data is not *tidy*.

To make this data tidy, we'll do some cleanup. First, remove the columns you don't need, using the verb `select`.

HINT: `?select` and the examples of helper functions will help you be efficient.

```{r select}
d.tidy <- select(...)
```


Try renaming some variables using `rename`. A good naming scheme is:

* consistent with case
* consistent with "." or "_" ( "_" is usually preferred)
* concise as will be comprehensible to others

Try using the `%>%` operator as well. So you will be "piping" `d %>% rename(...)`.

```{r rename}
d.tidy <- ...
```


OK, now for the tricky part. Use the verb *gather* to turn this into a *tidy* data frame.

HINT: look for online examples!

```{r gather}
d.tidy <- ...
```


Bonus problem: *spread* these data back into a wide format data frame.

```{r spread}
d.wide <- ...
```


Part 3: Manipulating the data using dplyr
-----------------------------------------

Try also using the dplyr `distinct` function to remove the duplicate participants from the raw csv file that you discovered in part 1.

```{r}
d.raw <- read.csv("../data/janiszewski_rep_exercise.csv")
d.unique.subs <- ...
```

