---
title: 'Psych 254 W15 PS #2'
author: "Nicholas Moores"
date: "February 4, 2015"
output: html_document
---

Give a hint of what they're looking at with the behavioral data and what they found
then go into imaging 
ROI method
ROI data
Searchlight method
Searchlight data
etc.

Results are a little confusing -- they found some nice evidence for the hippocampus stuff but the rest seems exploratory

Background reading: 
Polyn et al, 2009 computational model is dense
Haxby 2001 --> similar methods (grandfather but not the great grandfather)
Make sure you read the supplement -- if there's relevant stuff in there have back up slides. Have the supplement available.

If people ask about data modeling or slice prescription just say that 'I'm still learning about neuroimaging. I can tell you what the words say but I don't quite know what it means yet.' Focus on the pattern similarity analysis section!

Modeled every trial on its own; trying to get an estimate of brain activity for each image on the screen. Pattern Similarity Analysis: for each of the encoding trials, get an estimate of what the activity int he brain looked like during each trial. For each voxel, for each trial, there was a number representing what they think the brain response was. For two trials, they take all the voxels for one trial and put them into a vector, take all the voxels for the other trial and put them into a vector, then take the correlation between them. If you get high correlation, that's high pattern similarity

Pattern similarity is a common enough analysis that people will know what it means; treating your 3D cube of values as a single vector, one to one correspondence for the same voxels in a vector representing the other trial, then correlation between the two.

This is problem set #2, in which we hope you will practice the visualization package ggplot2, as well as hone your knowledge of the packages tidyr and dplyr. 

Part 1: Basic intro to ggplot
=============================

Part 1A: Exploring ggplot2 using qplot
--------------------------------------

Note, that this example is from the_grammar.R on http://had.co.nz/ggplot2 
I've adapted this for psych 254 purposes

First install and load the package.

```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
```

Now we're going to use qplot. qplot is the easy interface, meant to replace plot. You can give it simple `qplot(x,y)` examples, or slightly more complex examples like `qplot(x, y, col=grp, data=d)`. 

We're going to be using the diamonds dataset. This is a set of measurements of diamonds, along with their price etc.

```{r}
head(diamonds)
qplot(diamonds$carat, diamonds$price)
```

Scatter plots are trivial, and easy to add features to. Modify this plot so that it uses the dataframe rather than working from variables in the general namespace (good to get away from retyping `diamonds$` every time you reference a variable). 

```{r}
ggplot(diamonds, aes(x=carat, y=price)) +
  ggtitle("Diamonds by Carat and Price") +
  xlab("Diamond Carats") + ylab("Diamond Price") + 
    geom_point(shape=16, 
               color = 'black',
               alpha = 0.5)
```

Try adding clarity and cut, using shape and color as your visual variables. 

```{r}
ggplot(diamonds, aes(x=carat, y=price)) +
  ggtitle("Diamonds by Carat and Price") +
  xlab("Diamond Carats") + ylab("Diamond Price") + 
    geom_point(aes(shape=clarity, 
               color = cut))
```
One of the primary benefits of `ggplot2` is the use of facets - also known as small multiples in the Tufte vocabulary. That last plot was probably hard to read. Facets could make it better. Try adding a `facets = x ~ y` argument. `x ~ y` means row facets are by x, column facets by y. 

```{r}
ggplot(diamonds, aes(x=carat, y=price)) +
  ggtitle("Diamonds by Carat and Price") +
  xlab("Diamond Carats") + ylab("Diamond Price") + 
    geom_point(aes(shape=clarity, 
               color = cut)) +
  facet_grid(clarity ~ cut)
```

But facets can also get overwhelming. Try to strike a good balance between color, shape, and faceting.

HINT: `facets = . ~ x` puts x on the columns, but `facets = ~ x` (no dot) *wraps* the facets. These are underlying calls to different functions, `facet_wrap` (no dot) and `facet_grid` (two arguments). 

```{r}
```

The basic unit of a ggplot plot is a "geom" - a mapping between data (via an "aesthetic") and a particular geometric configuration on coordinate axes. 

Let's try some other geoms and manipulate their parameters. First, try a histogram (`geom="hist"`). 

```{r}
```

Now facet your histogram by clarity and cut. 

```{r}
```

I like a slightly cleaner look to my plots. Luckily, ggplot allows you to add "themes" to your plots. Try doing the same plot but adding `+ theme_bw()` or `+ theme_classic()`. Different themes work better for different applications, in my experience. 

```{r}
```

Part 1B: Exploring ggplot2 using ggplot
---------------------------------------

`ggplot` is just a way of building `qplot` calls up more systematically. It's
sometimes easier to use and sometimes a bit more complicated. What I want to show off here is the functionality of being able to build up complex plots with multiple elements. You can actually do this using qplot pretty easily, but there are a few things that are hard to do. 

`ggplot` is the basic call, where you specify A) a dataframe and B) an aesthetic mapping from variables in the plot space to variables in the dataset. 

```{r}
d <- ggplot(diamonds, aes(x=carat, y=price)) # first you set the aesthetic and dataset
d + geom_point() # then you add geoms
d + geom_point(aes(colour = carat)) # and you can keep doing this to add layers to the plot
```

Try writing this as a single set of additions (e.g. one line of R code, though you can put in linebreaks). This is the most common workflow for me. 


```{r}
```


You can also set the aesthetic separately for each geom, and make some great plots this way. Though this can get complicated. Try using `ggplot` to build a histogram of prices. 

```{r}
```

Part 2: Diving into real data: Sklar et al. (2012)
==================================================

Sklar et al. (2012) claims evidence for unconscious arithmetic processing. We're going to do a reanalysis of their Experiment 6, which is the primary piece of evidence for that claim. The data are generously contributed by Asael Sklar. 

First let's set up a few preliminaries. 

```{r}
library(tidyr)
library(dplyr)

sem <- function(x) {sd(x) / sqrt(length(x))}
ci95 <- function(x) {sem(x) * 1.96}
```

Data Prep
---------

First read in two data files and subject info. A and B refer to different trial order counterbalances. 

```{r}
subinfo <- read.csv("../data/sklar_expt6_subinfo_corrected.csv")
d.a <- read.csv("../data/sklar_expt6a_corrected.csv")
d.b <- read.csv("../data/sklar_expt6b_corrected.csv")
```

Gather these datasets into long form and get rid of the Xs in the headers.

```{r}
subinfo <- tbl_df(subinfo)
d.a <- tbl_df(d.a)
d.b <- tbl_df(d.b)
d.a.tidy <- d.a %>% gather("x_type", "RT", X1:X21, na.rm=TRUE) %>% mutate(x_type = gsub("X","",x_type))
d.b.tidy <- d.b %>% gather("x_type", "RT", X22:X42, na.rm=TRUE) %>% mutate(x_type = gsub("X","",x_type))
glimpse(subinfo)
glimpse(d.a.tidy)
glimpse(d.b.tidy)
```

Bind these together. Check out `bind_rows`.

```{r}
d.all.tidy <- bind_rows(d.a.tidy, d.b.tidy)
```

Merge these with subject info. You will need to look into merge and its relatives, `left_join` and `right_join`. Call this dataframe `d`, by convention. 

```{r}
subinfo <- subinfo %>% rename(subjectID = subid)
d.all.tidy <- d.all.tidy %>% rename(subjectID = x_type)
subinfo$subjectID <- as.factor(subinfo$subjectID)
d.all.tidy$subjectID <- as.factor(d.all.tidy$subjectID)
d.all.tidy <- left_join(subinfo, d.all.tidy, by="subjectID")
d.all.tidy$subjectID <- as.factor(d.all.tidy$subjectID)
```

Clean up the factor structure.

```{r}
d.all.tidy$presentation.time <- factor(d.all.tidy$presentation.time)
levels(d.all.tidy$operand) <- c("addition","subtraction")
```

Data Analysis Preliminaries
---------------------------

Operand x congruency x presentation time. Challenging to know what to collapse and what not to.

Examine the basic properties of the dataset. First, take a histogram.

```{r}
m.rt <- exp(mean(log(d.all.tidy$RT),na.rm=TRUE)) #take log average of RT: average natural logs of RTs, then take the exponential of that
sd.rt <- exp(sd(log(d.all.tidy$RT),na.rm=TRUE)) #geometric standard deviation: exponentiated value of the standard deviation of the log-transformed values
qplot(RT, data=d.all.tidy) + 
  geom_vline(xintercept=m.rt + 2*sd.rt, col="red", lty=2) + #lets us view how much is spread 2 SDs above mean
  geom_vline(xintercept=m.rt - 2*sd.rt, col="red", lty=2) + #lets us view how much is spread 2 SDs below mean
  xlab("RT spread") +
  ylab("Frequency") +
  ggtitle("RT Histogram of Sklav Data (marking +/- 2 SDs)")
```

Challenge question: what is the sample rate of the input device they are using to gather RTs?

```{r}
```

Sklar et al. did two manipulation checks. Subjective - asking participants whether they saw the primes - and objective - asking them to report the parity of the primes (even or odd) to find out if they could actually read the primes when they tried. Examine both the unconscious and conscious manipulation checks (this information is stored in subinfo). What do you see? Are they related to one another?

Run a logistic regression on it! They are related. Put a red box around people that they kept.
```{r} 
```

OK, let's turn back to the measure and implement Sklar et al.'s exclusion criterion. You need to have said you couldn't see (subjective test) and also be not significantly above chance on the objective test (< .6 correct). Call your new data frame `ds`.

```{r}
```

Sklar et al.'s analysis
-----------------------

Sklar et al. show a plot of a "facilitation effect" - the time to respond to incongruent primes minus the time to respond to congruent primes. They then show plot this difference score for the subtraction condition and for the two presentation times they tested. Try to reproduce this analysis.

HINT: first take averages within subjects, then compute your error bars across participants, using the `sem` function (defined above). 
somewhere: do objective.test x facilitation scatterplot with lm

somwhere: do distance x RT scatterplot, with color by congruent
```{r}
```

Now plot this summary, giving more or less the bar plot that Sklar et al. gave (though I would keep operation as a variable here. Make sure you get some error bars on there (e.g. `geom_errorbar` or `geom_linerange`). 

```{r}
```

What do you see here? How close is it to what Sklar et al. report? Do the error bars match? How do you interpret these data? 

Challenge problem: verify Sklar et al.'s claim about the relationship between RT and the objective manipulation check.

```{r}
```

Your own analysis
-----------------

Show us what you would do with these data, operating from first principles. What's the fairest plot showing a test of Sklar et al.'s original hypothesis that people can do arithmetic "non-consciously"?

```{r}
```

Challenge problem: Do you find any statistical support for Sklar et al.'s findings?

```{r}
```
